### 2018-10-22 eurocg/short_abstract/abstract.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%% article.tex %%%%%%%%%%%%%%%%%%%%%%%
%
% edit this file for the abstract of your own talk
% and save it under the original name "abstract.tex".
% Then run "pdflatex eurocg18-short-abstract.tex" to see how
% your abstract will look like.
%

\authors{Michael H\"oller, Fabian Klute, Soeren Nickel, Martin~N\"ollenburg, \speaker{Birgit Schreiber}}
% The author who will give the presentation should be marked with the
% "\speaker{...}" command

\title{Maximizing Ink in Symmetric Partial Edge Drawings of $k$-plane Graphs}

% Exactly one of the following three lines should be present
%

%\nofigures
\onefigure{MISPE-fig} % In this case, we have one thumbnail figure
%\twofigures{SIRAR-fig1}{SIRAR-fig2}

% To distinguish your figures from those of other particupants,
% the file names should start with the first five initials of the
% title. (SIRAR in the example)
%
% The figures must be PDF files with the extension .pdf
% (SIRAR-fig1.pdf in the example), and they must reside in the same
% directory as the main file.
%
% The figures will be scaled to a height of 25 millimeters and
% enclosed by boxes with a margin of 3pt. It is best to design your
% figures in this size. Otherwise, you should make sure that the
% thickness of your lines survives the rescaling.

% By default, the figures will be enclosed in a box.
% If your figures already have an encloding box, you can disable the
% addition of an enclosing box by uncommenting the following line:
%
% \FigureBorderfalse

\begin{abstract}
  % Here comes the text of your abstract.
  % A good length is 6-15 lines in the output. The limit is 20 lines.
  % The presence of thumbnail pictures reduces the available amount
  % for text.
%
	We study symmetric partial edge drawings (SPEDs) for non-planar graphs, in which edges are drawn only partially as pairs of opposing stubs of equal length on the respective end-vertices. It is known that maximizing the ink (or the total stub length) when transforming a straight-line drawing with crossings into a SPED is tractable for 2-plane input drawings, but generally NP-hard. We show that the problem remains NP-hard even for 3-plane input drawings. We also present efficient algorithms for ink maximization of $k$-plane input drawings whose edge intersection graph forms a collection of trees or cacti. %\verb|\FigureBorderfalse|.

\end{abstract}

%%%%%%%%%% =========================================